---
title: "ECON6027_1b"
author: "Dr. Liu"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install sf

If you are installing sf (or any other package for the first time, use the command "install.packages('sf', dep=T)" or the "Install" button on the side panel.

Troubleshooting and more info: <https://r-spatial.github.io/sf/>

```{r}
is.element('sf', installed.packages())
library(sf)
```

At the time of updating my codes, the most recent version of GEOS, GDAL, and PROJ are: GEOS 3.12.2, GDAL 3.9.2, PROJ 9.4.1

## Read a shapefile to R

Remember to change the working directory first. The command below reads a MULTIPOLYGON simple feature collection with 323 features and 15 fields

```{r}
sg_pa = st_read("MP14_SUBZONE_WEB_PL.shp") 
```

## Inspect the shapefile

```{r}
class(sg_pa)
summary(sg_pa)
head(sg_pa)
table(st_is_valid(sg_pa))
```

There are 9 false entries. As a habit you should always check the validity especially if the dataset is downloaded from the Internet.

If there are FALSE entries, use the command st_make_valid() to validate the object.

```{r}
sg_pa = st_make_valid(sg_pa)
table(st_is_valid(sg_pa))
```

No more invalid entries!

## Visualise the shapefile

```{r}
plot(sg_pa, max.plot = 12)
plot(sg_pa["PLN_AREA_N"])
```

## Save an sf object to the hard drive

In the st_write() function, you can use the "dsn" (data source name) argument to specify which driver or the file format you need. For example "xxx.shp" will write the sf object to a shapefile as demonstrated below.

```{r}
st_write(sg_pa, "MySingapura.shp", delete_layer=T) 
```

The "delete_layer = TRUE" argument will overwrite any other file of the same name.

If instead you want to save as a geojson file, use the following command.

```{r}
#st_write(sg_pa, "MySingapura.geojson")
```
