---
title: "ECON6027_1c"
author: "Dr. Liu"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Create a simple data.frame

Import the Excel data to the R console. There are many ways to do this. My personal favourite for small datasets like this is to import from the clipboard. You should copy the Excel data to the clipboard before executing the following command.

```{r}
fav = read.csv("1c_data.csv")
class(fav)
plot(fav)
```

We have a spatial dataset, but this dataset has no “spatial awareness”. Let’s give it some spatial powers!

The data.frame will now be converted to a sf object using the sf package by identifying the coordinates as follows.

```{r}
library(sf)
(fav.sf = st_as_sf(fav, coords=c("lon_x","lat_y")))
plot(st_geometry(fav.sf), main="Bucket List")
```

The “coords” argument follows the cartesian coordinates convention (x,y). Notice the lack of CRS: R knows these are locations but not quite sure where...

## Roll up your sleeves

Load the "world" dataset from spData. Install this package if you have not done so already. spData is a diverse spatial datasets for demonstrating, benchmarking and teaching spatial data analysis.

We will load and inspect the "world" dataset.

```{r}
world = st_read(system.file("shapes/world.gpkg", package="spData"))
world
```

Now, let us inspect the dataset.

```{r}
table(st_is_valid(world))
class(world)
dim(world) # 177 features, 11 attributes
names(world)
summary(world)
head(world)
```

Geometry column of the dataset contains the coordinate data. If you remove the geometry column, the sf object reverts to a simple data.frame.

```{r}
world$geom
world_df = st_drop_geometry(world)
class(world_df)
head(world_df)
```

## Initial plots

```{r}
# plot(world)
plot(world[3:6])
```

Basic plot of the countries and continents based on column name.

```{r}
plot(world[c("name_long", "continent")]) 
```

## Spatial dataset manipulation

In this section we will see some important ways to manipulate a spatial dataset. Most of these commands can be applied to any dataset in general.

### 1. Subsetting using base R functions [] or subset()

The following code chunk shows how to subset based on columns.

```{r}
world[, 7:10] # Extract columns 7 to 10
summary(world[, 7:10]) # numerical summary of the columns
```

The following code chunk allows subsetting based on certain criterion: lad area more than 5 million square km.

Method 1

```{r}
l_area = world$area_km2 > 5000000
class(l_area)
table(l_area)
(big_countries = world[l_area, ])

```

Method 2

```{r}
(big_countries = subset(world, area_km2 > 5000000)) # method 2
```

Since Antartica is not a country, we will remove it from the dataset.

```{r}
(big_countries = big_countries[-7, ]) # remove Antarctica
```

In the next chunk, we will extract and highlight Asia (a specific geographic location) in a map. 

Using [] is one way to subset a dataset. Notice the resulting dataset is also an sf object.

```{r}
table(world$continent)
asia = world[world$continent == "Asia", ] 
plot(world["continent"], reset = F) 
plot(asia, add = T, col = "black")
```

Plot Sri Lanka (Also a possible solution to Exercise B)

```{r}
world$name_long 
(sl = world[world$name_long == "Sri Lanka", ])
plot(st_geometry(asia), main="Asia")
plot(st_geometry(sl), col = "red", add=T)
```

plot(st_geometry()) gives an outline of the geometry column.

### 2. Attribute data aggregation

Aggregation operations summarise datasets by a ‘grouping variable’, typically an attribute column. In the following example, we aggregate the population by continent.

```{r}
(cont_pop = aggregate(pop ~ continent, FUN = sum, data = world, na.rm = T)); class(cont_pop) # output is a data.frame
(cont_pop2 = aggregate(world['pop'], by = list(world$continent), FUN = sum, na.rm = T)); class(cont_pop2) # results in an sf output
# (cont_pop2 = aggregate(world$pop, by = list(world$continent), FUN = sum, na.rm = T)); class(cont_pop2)
```

### 3. Attribute joining using "key variable" 

Idea is to join data from different sources. This is useful when you want to “give” spatial awareness to a dataset. In the following example, we will use the "coffee_data" dataset from spData and join that to the "world" dataset using the key variable "name_long"

```{r}
library(spData)
summary(coffee_data)
class(coffee_data)
library(dplyr)
# left join
(world_coffee = left_join(world, coffee_data))
plot(world_coffee["coffee_production_2016"])
# inner join
(world_coffee2 = inner_join(world, coffee_data)) 
```

Notice the joining is done using the shared “key” variable “name_long”. By default,dplyr uses all variables with matching names. In this case, both world_coffee and world objects contained a variable called name_long, explaining the message Joining, by = "name_long". Where variable names are not the same, you have two options: (1) Rename the key variable in one of the objects so they match, or (2) Use the by argument to specify the joining variables.

The difference between the two functions is that left_join() fully preserves the features of the left dataset while inner_join() will find the common elements belonging to both datasets. You may choose the desired function based on the final dataset you need.

### 4. Create new attributes

```{r}
world_coffee$prod_yoy = (world_coffee$coffee_production_2017/world_coffee$coffee_production_2016 - 1)*100
world_coffee
```

