---
title: "ECON6027_4b_updated"
author: "Dr. Liu"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spatial desctiptive summary measures

Start by loading the packages:

```{r}
library(aspace, quietly = T); library(sf)
```

### Datasets

Change working directory and load datasets. The datasets are the Singapore planning area shapefile and three more shapefiles containing the locations of the speed camera in Singapore.

```{r}
sg_regions = st_read("MP14_PLNG_AREA_WEB_PL.shp") # sg regions
table(st_is_valid(sg_regions)) # Validity must be checked for all loaded polygons
sg_regions = st_make_valid(sg_regions); table(st_is_valid(sg_regions))

SLC = st_read("SPF_PSLC.shp") # speed laser cameras
table(st_is_valid(SLC)) # usually points objects have valid geometry
```

CRS matters: Before we conduct summary calculations, the unit of analysis must be in a metric form so that distance calculations will be meaningful. If the dataset follows a geographic CRS (st_is_longlat() is TRUE), then you must transform the CRS to a projected CRS.

```{r}
st_crs(sg_regions)$proj
st_crs(SLC)$proj
isTRUE(all.equal(st_crs(sg_regions), st_crs(SLC)))
```

Both are projected with the same CRS. So, we are good to go!

### Initial maps

Start by creating a basemap "sg_map" and add the laser camera locations.

```{r}
library(tmap)
(sg_map = tm_shape(sg_regions) + tm_polygons(col="REGION_N", alpha=0.5, 
                          border.col = "grey", palette="Set2", legend.show = F))
(LC_locations = sg_map + 
    tm_shape(SLC) + tm_dots(size=0.1) + 
    tm_layout(main.title="Locations of Laser Speed Cameras, Singapore", 
              inner.margins = 0.1))
```

### Descriptive summary measures

Computations using the aspace package functions can only be carried out on cartesian coordinates given in matrix form. Thus we will extract the coordinates of the speed laser camera locations of SLC.

```{r}
LC_coords = st_coordinates(SLC)
class(LC_coords) 
```

#### Measures of central tendency

Compute the mean and the median and assign it to a list object for later use.

```{r}
mean = calc_mnc(id=1, points=LC_coords)
mean$LOCATIONS
median = calc_mdc(id=2, points= LC_coords)
median$LOCATIONS
```

#### SDD (standard distance disc)

In the following chunk, we compute the standard distance and the output is assigned to a list object called "sdd".

```{r}
sdd = calc_sdd(id=3, calccentre = T, weighted=F, points = LC_coords)
```

In order to plot the standard distance disc using tmap, we need to convert standard distance disc contained in the list "sdd$LOCATIONS" created in the previous chunk to a sf line object.

```{r}
ls()
(LC_sdd_line = sdd$LOCATIONS %>% st_as_sf(coords = c("x", "y"), crs=st_crs(SLC))
    %>% st_combine() %>% st_cast("LINESTRING")) # CRS from the parent object.
(LC_std_dist = st_sf(sdd$ATTRIBUTES, geom=st_geometry(LC_sdd_line))) 
# create sf object by adding in the attribute data
```

#### SDE (standard deviational ellipse)

Coding to generate the SDE is much similar to that of SDD above.

```{r}
sde = calc_sde(id=4, calccentre = T, weighted=F, points= LC_coords)
```

As before we convert standard deviational ellipse to sf line object to help with tmap mapping

```{r}
ls()
(LC_sde_line = sde$LOCATIONS %>% st_as_sf(coords = c("x", "y"), crs=st_crs(SLC)) 
  %>% st_combine() %>% st_cast("LINESTRING"))
(LC_std_ellps = st_sf(sde$ATTRIBUTES, geom=st_geometry(LC_sde_line)))
```

#### Plot of the summary measures

Create a point object with mean and median to be plotted as a separate layer. SDD and SDE will be added as additional layers. (Note the three different tm_shape layers).

```{r}
mean$LOCATIONS; median$LOCATIONS
(mean_median = rbind(mean$LOCATIONS, median$LOCATIONS) %>% data.frame() %>% 
  st_as_sf(coords=c("x","y"), crs=st_crs(SLC))) # matrix -> data.frame -> sf points
mean_median$id = c("Spatial Mean","Spatial Median") # assign names
mean_median
LC_locations + 
  tm_shape(LC_std_ellps) + tm_lines(lwd=2) + 
  tm_shape(LC_std_dist) + tm_lines(lwd=2, lty=2) + 
  tm_shape(mean_median) + tm_dots(size=0.1, shape="id", col="red", title.shape="Measures") + 
  tm_compass() +
  tm_add_legend(type="line", labels="SDD", lty=2) + 
  tm_add_legend(type="line", labels="SDE", lty=1)
```

### Weighted descriptive summary measures

We will use the dengue clusters as of May 2020 dataset for this analysis. The centroid of the clusters will be analysed. The weights are assigned based on the size of the cluster.

```{r}
dengue = st_read("dengue-clusters-may20.geojson") # original CRS is WGS84
st_is_longlat(dengue)
dengue = st_transform(dengue, crs=st_crs(sg_regions)) # align the CRS
isTRUE(all.equal(st_crs(sg_regions), st_crs(dengue)))

# Also check the validity of the geometry
table(st_is_valid(dengue)) 
dengue = st_make_valid(dengue); table(st_is_valid(dengue)) 
```

In the following chunk, we plot the dengue clusters on a map of Singapore.

```{r}
names(dengue)
dengue$area = st_area(dengue) # new attribute which will be used as weights later
summary(dengue$area)
sg_map + 
  tm_shape(dengue) + tm_polygons(col="area", title="Cluster size") + 
  tm_layout(main.title = "Dengue clusters, May 2020", legend.outside = T)
```

For the purpose of computing summary measures, extract the coordinates of the center of these clusters.

```{r}
dengue_spots = st_centroid(dengue)
dengue_coords = st_coordinates(dengue_spots) # extract coordinates from dengue_spots
(dengue_map = sg_map + 
  tm_shape(dengue_spots) + tm_dots() +
  tm_layout(main.title="Locations of High Dengue Incidence, May 2020, Singapore", 
            inner.margins = 0.1))
```

The following code chunk demonstrates how to extract weights. Note that aspace requires the coordinates to be unit free. However, ”dengue_area” is in square meters which we will remove now. (if your weights are in frequencies, you don’t have a problem.)

```{r}
(dengue_area = st_area(dengue))
class(dengue_area)
(dengue_area = units::drop_units(st_area(dengue)))
class(dengue_area)
```

The measures are computed as follows (Note that the objects created in the previous section are being overwritten. If you want to create a new set of objects, change the name of the assignment):

```{r}
mean = calc_mnc(id=1, weighted = T, weights = dengue_area, points = dengue_coords)
sdd = calc_sdd(id=2, calccentre = T, weighted = T, weights = dengue_area, points = dengue_coords)
sde = calc_sde(id=3, calccentre = T, weighted = T, weights = dengue_area, points = dengue_coords)
```
