---
title: "ECON6027_6a"
author: "Dr. Liu"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read neighbours from an existing GAL file 

Load the required libraries:

```{r}
library(sf); library(spdep)
```

Contiguity neighbour GAL file for UK

```{r}
(UK_nb = read.gal("uk_cont.gal"))
class(UK_nb)
```

nb is a neighbour list object which is how neighbourhood information is handled in the R ecosystem.

We also need a shapefile to anchor the external “nb” object to…

```{r}
(UK = st_read("NUTS_Level_1__January_2018__Boundaries.shp"))
table(st_is_valid(UK)) # all valid
```

In general, after this step, you need to make sure the areal units are in the same order as the “nb” object. For your convenience, in this case, I have made sure that the object ID of the shapefile and the gal file are the same.

Next, we will plot the contiguity relationships among the regions of UK on a map of UK.

```{r}
UK_coords = st_point_on_surface(st_geometry(UK))
plot(st_geometry(UK), border="grey60", reset = F, main="nb from GAL file") 
plot(UK_nb, UK_coords, pch=19, cex=0.6, add=T)
# write.nb.gal(UK_nb, "UK_nb_new.gal", oldstyle=F) # create GAL file
```

See ?plot.nb for more details on plotting nb objects.

## Create neighbours based on spatial polygons

We will work with Singapore planning areas.

```{r}
SG = st_read("MySingapura.shp")
table(st_is_valid(SG))
SG = st_make_valid(SG); table(st_is_valid(SG))
```

### 1. Contiguous neighbours

The poly2nb() function can be used to convert polygon objects to nb objects. The default is queen contiguity neighbours from SG. If we make queen=F, then we get the rook contiguity.

```{r}
(SG1_nb = poly2nb(SG)) # queen contiguity
# write.nb.gal(SG1_nb, "SG1_nb.gal", oldstyle=F)
(SG2_nb = poly2nb(SG, queen=F)) # rook contiguity
# check the snap argument
```

Let us plot the contiguity neighbours on a map of Singapore.

```{r}
SG_coords = st_point_on_surface(st_geometry(SG))

plot(st_geometry(SG), border="grey60", axes=T,  main="Queen Contiguity")
plot(SG1_nb, SG_coords, pch=19, cex=0.6, add=T)

plot(st_geometry(SG), border="grey60", axes=T,  main="Rook Contiguity")
plot(SG2_nb, SG_coords, pch=19, cex=0.6, add=T)

isTRUE(all.equal(SG1_nb, SG2_nb, check.attributes = F))
```

For Singapore planning areas queen contiguity is the same as rook contiguity.

### 2. k nearest neighbours

```{r}
(IDs = row.names(SG))
SG3_nb = knn2nb(knearneigh(SG_coords, k = 1), row.names = IDs) # nearest neighbour
SG4_nb = knn2nb(knearneigh(SG_coords, k = 2), row.names = IDs) # 2 nearest neighbours
SG5_nb = knn2nb(knearneigh(SG_coords, k = 3), row.names = IDs) # 3 nearest neighbours
```

Note that metric distances may vary if st_centroid is used. Let us plot the k nearest neighbours on a map of Singapore.

```{r}
plot(st_geometry(SG), border="grey60", axes=T,  main="k=1")
plot(SG3_nb, SG_coords, pch=19, cex=0.6, add=T)

plot(st_geometry(SG), border="grey60", axes=T,  main="k=2")
plot(SG4_nb, SG_coords, pch=19, cex=0.6, add=T)

plot(st_geometry(SG), border="grey60", axes=T,  main="k=3")
plot(SG5_nb, SG_coords, pch=19, cex=0.6, add=T)

plot(st_geometry(SG), border="grey60", axes=T,  main="Compare: k=2 vs. k=3 (difference in red)")
plot(SG5_nb, SG_coords, col="red", pch=19, cex=0.6, add=T) # plot the nb with more connections first.
plot(SG4_nb, SG_coords, pch=19, cex=0.6, add=T)
```

The final plot shows the difference between the neghbourhood relationships when k=2 and k=3. 

### 3. Distance based neighbours

The following chunk create nb objects based on three distances and plots them.

```{r}
st_crs(SG)$units
SG6_nb = dnearneigh(SG_coords, d1 = 0, d2 = 1000, row.names = IDs)
SG7_nb = dnearneigh(SG_coords, d1 = 0, d2 = 5000, row.names = IDs)
SG8_nb = dnearneigh(SG_coords, d1 = 0, d2 = 10000, row.names = IDs)

plot(st_geometry(SG), border="grey60", axes=T,  main="Within 1km")
plot(SG6_nb, SG_coords, pch=19, cex=0.6, add=T)

plot(st_geometry(SG), border="grey60", axes=T,  main="Within 5km")
plot(SG7_nb, SG_coords, pch=19, cex=0.6, add=T)

plot(st_geometry(SG), border="grey60", axes=T,  main="Within 10km")
plot(SG8_nb, SG_coords, pch=19, cex=0.6, add=T)
```

As can be seen from the plots, note that distance-based neighbours can leave some spatial units neighbour-less and the degree of connectedness can exponentially increase as you increase the distance. 

## Create a weights list object from nb object

The nb list object that we created before contains information on how areal units are related to one another. It must be converted to a weights matrix before it can be used in statistical analyses. We can do this by suing the function nb2listw(). This function converts the nb list to a weights list object that contains information about the type of the weights matrix to be used in the analysis. Example:

```{r}
(UK_lw = nb2listw(UK_nb)) # row normalised weights
names(UK_lw)
names(attributes(UK_lw))
summary(unlist(UK_lw$weights))
summary(sapply(UK_lw$weights, sum))
```

There are many style options (see ?nb2listw). The default is the row normalised form. Row standardising can turn a symmetric weights matrix into an asymmetric one. 
There is also the chance of boosting the weights of the units near the boundary. 

In many econometric applications it is the norm to use style=“W” matrix as it has the property of sum(W)=n which is quite useful in numerical optimisations and other computations related to estimation of parameters and test statistics. (See next chapter).

By default, zero.policy=F, which means there cannot be “loners”. You can disable this by setting zer.policy=T but in many econometric applications such a W matrix raises more problems than can be answered, for example, the effective sample size, a.k.a degrees of freedom in tests model identification issues, etc. Example:

```{r}
# (SG1_lw = nb2listw(SG1_nb))
# (SG1_lw = nb2listw(SG1_nb, zero.policy=T))
# print(SG1_lw, zero.policy=T)
```


