---
title: "ECON6027_7a_sols"
author: "Dr. Liu"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solutions to ECON6027_7a

```{r}
library(spdep, quietly = T); library(sf, quietly = T)
```

Data prep is done first. The variables of the UK dataset are:

GVA: gross value added

LP: labour productivity

PBB: percentage of new business births

```{r}
UK_nb = read.gal("UK_nb.gal") #read GAL from the last chapter
(W_list = nb2listw(UK_nb))
uk_data = st_read("uk_data.shp")
names(uk_data)
st_is_valid(uk_data)
options(scipen = 999) # to remove “e” notation from the summary outputs
# options(scipen = 0) # to resore “e” notation 
```

### Solutions A

- Model 1:

```{r}
model1 = lm(formula = gross_v ~ L_prod + pct_bsb, data=uk_data)
summary(model1)
```

Conclusion: pct_bsb is not significant, but L_prod is significant.

- Model 2:

```{r}
model2 = lm(formula = gross_v ~ L_prod, data=uk_data)
summary(model2)
```

Conclusion: L_prod is significant.

- Model 3:

```{r}
model3 = lm(formula = gross_v ~ pct_bsb, data=uk_data)
summary(model3)
```

Conclusion: pct_bsb is significant, possibly due to omitted variable bias.

So, which model should we choose? Model 2 as it has the largest adjusted R-square statistic (0.8899) and the most significant F-statistic (p-value: 0.000002708).

Alternatively, you can also consider AIC and BIC to compare.

```{r}
AIC(model1);AIC(model2);AIC(model3)
BIC(model1);BIC(model2);BIC(model3)
```

Conclusion: model 2 has the lowest AIC and BIC.

- Model 4

```{r}
model4 = lm(formula = L_prod ~ pct_bsb, data=uk_data)
summary(model4)
```

Conclusion: Given the significance of model 4, there maybe endogeneity in model 1.

#### Hypothesis tests

- Breusch–Pagan test

```{r}
lmtest::bptest(model1)
```

Conclusion: Null of homoskedasticity is not rejected. No evidence of heteroskedasticity.

- Jarque–Bera test

```{r}
tseries::jarque.bera.test(model1$residuals)
shapiro.test(model1$residuals)
```

Conclusion: Null of normality is not rejected. 

### Solutions B

Lagged labour productivity variable (W*L_prod)

```{r}
as.matrix(lag.listw(W_list,uk_data$L_prod)) # lagged L_prod
```

### Solutions C

We consider model 1

```{r}
(L_GVA = as.matrix(lag.listw(W_list,uk_data$gross_v))) # lagged gross_v
moran.plot(uk_data$gross_v, W_list)
```

There seem to be positive spatial autocorrelation in the dependent variable “gross_v” with most points in the H-H (hotspots) and L-L quadrants (coldspots).

```{r}
moran.test(uk_data$gross_v, W_list)
```

Conclusion: no evidence of spatial autocorrelation in the dependent variable at 5% level.

```{r}
moran.plot(model1$residuals, W_list)
lm.morantest(model1, W_list, alternative="less") # Moran's test for regression residuals
```

Conclusion: Do not reject H0. No evidence of spatial autocorrelation in the OLS residuals.


