---
title: "ECON6027_9a"
author: "Dr. Liu"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RASTER DATA ANALYSIS

## Introduction

In contrast to the vector data model  (which represents points, lines and polygons as discrete entities in space), raster data usually represent continuous surfaces such as elevation, temperature, etc. Of course, we can represent discrete features such as soil or land-cover classes also with the help of a raster data model.

The greographic raser data model consists of:

- A raster header and

- A matrix representing equally spaced cells (pixels). This matrix representation avoids storing explicitly the coordinates for the four corner points (in fact it only stores one coordinate, namely the origin) of each cell corner as would be the case for rectangular vector polygons. This and map algebra makes raster processing much more efficient and faster than vector data processing. 

The raster header defines:

1. The CRS

2. The extent; the number of columns, the number of rows and the cell size resolution

3. The origin; starting from the origin, we can easily access and modify each single cell by either using the ID of a cell or by explicitly specifying the rows and columns. 

## Raster Data in R

Popular raster data file formats in R are:

1. GeoTiff: Popular raster format. A TIFF file containing additional spatial metadata.

2. File Geodatabase

3. Arc ASCII: Text format where the first six lines represent the raster header, followed by the raster cell values arranged in rows and columns.

In R, the popular packages you can used to handle raster data are (1) terra and (2) stars. The former is built on its predecessor, the "raster" package. We will use the "terra" package.

```{r}
library(terra)

# install.packages("spDataLarge", repos = "https://geocompr.r-universe.dev") # install the required package
```

A quick illustration of a typical raster dataset is given below using a tiff dataset in "spDataLarge".

```{r}
raster_filepath = system.file("raster/srtm.tif", package = "spDataLarge") # identify the file
zion = rast(raster_filepath) # rasterise the tif
class(zion) # Notice the new class “Spatraster”.
zion # shows the raster header

# Basic maps
plot(zion, main="Zion National Park")
```

### Create a simple raster from scratch

We start by creating a default raster.

```{r}
rast() # default empty raster.
```

Note the following default options:

- The new class “Spatraster”.

- Dimensions to cover 180 latitudes and 360 longitudes.

- Each pixel to be of unit length on all sides.

- CRS is WGS84.

Next, we deviate from the default options to create our own raster and also assign values to the empty cells (pixels).

```{r}
new_raster = rast(nrows = 6, ncols = 6, xmin = -1.5, xmax = 1.5, ymin = -1.5, ymax = 1.5, vals = 1:36)
new_raster2 = rast(nrows = 100, ncols = 100, xmin = -25, xmax = 25, ymin = -25, ymax = 25, vals = 1:10000)

# plot
par(mfrow=c(2,1)); plot(new_raster, main="Raster 1"); plot(new_raster2, main="Raster 2"); par(mfrow=c(1,1))

new_raster
# new_raster2
```

Given the number of rows and columns as well as the extent (xmin, xmax, ymin, ymax), the resolution must be 0.5. The unit of the resolution is that of the underlying CRS. Here, it is degrees, because the default CRS of raster objects is WGS84. However, one can specify any other CRS with the crs argument.

## Illustration

### Things to do

- We will download a satellite image file in .tiff form and work with it.

- Obtain terrain information from the NOAA website: https://www.noaa.gov 

- We will consider data from the global relief model from NOAA. A global relief model (global topography model or composite model), combines digital elevation model (DEM) data over land with digital bathymetry model (DBM) data over water-covered areas (oceans, lakes) to describe Earth's relief. A relief model thus shows how Earth's surface would look like in the absence of water or ice masses. For more information on the Global relief model and other sources of relief model data, visit, https://en.wikipedia.org/wiki/Global_relief_model

### Steps to download the data

1. https://ngdc.noaa.gov/mgg/global/global.html 

2. Choose your desired location. I choose Sri Lanka owing to its versatile terrain properties and it is an island, so I don’t have to worry about other countries coming in. Another good candidate is New Zealand or even Singapore (although is Singapore is mostly flat).

3. Go to “Extract Custom Grid”

4. Select dataset. “ice” recommended. This gives a grid of Earth's surface depicting the top of the Antarctic and Greenland ice sheets (1-minute resolution). A minute of arc, denoted by the symbol ′, is a unit of angular measurement equal to 1/60 of one degree. Since one degree is 1/360 of a turn  (or complete rotation), one minute of arc is 1/21600 of a turn.

5. Scroll to the desired location.

6. Select area of interest by rectangle and save the tiff (tagged image file format) image to your working directory.

```{r}
(sl_raster = rast("SL_terrain.tiff"))
plot(sl_raster) #???
```

What do you think is going on? The tiff gave a rster that includes the entire selected region, not just Sri Lanka. We will extract the land area later.

Now, there is a different consideration you need to make. If you take (0,0) to be the origin of the latitudes (-90,90) and longitudes (-180,180), then Sri Lanka, which is roughly located at 7.8731° N, 80.7718° E, is in the 1st quadrant (if one may say so...). Thus, if one was to draw outline of Sri Lanka clockwise, you need the longitudes and latitudes to go in the opposite directions. For example, starting from Point Pedro (which is the north most point in Sri Lanka), when going clockwise, longitudes will increase and the latitudes will decrease. We will sort the data according to this requirement below.

```{r}
lons = seq(xmin(sl_raster), xmax(sl_raster), length.out = dim(sl_raster)[2])
lats = -sort(-seq(ymin(sl_raster), ymax(sl_raster), length.out = dim(sl_raster)[1])) # going in the opposite direction
sl_boundary = as.matrix(expand.grid(lons, lats))
sl_elev = c(t(as.matrix(sl_raster)))
head(sl_boundary); head(sl_elev)
```

Next we discard the excess terrain that does not belong to the land above sea level.

```{r}
sl_bb = maps::map.where("world", sl_boundary[ , 1], sl_boundary[ , 2])
table(sl_bb, useNA="ifany")
keep_cases = which((sl_bb == "Sri Lanka") | (sl_bb == "Sri Lanka:Jaffna") | (sl_bb == "Sri Lanka:Manar"))
sl_boundary = sl_boundary[keep_cases, ]
sl_elev = sl_elev[keep_cases]
```

In the following chunk, we convert the raster object to an sf object and make a quick plot.

```{r}
class(sl_boundary)
sl = data.frame(sl_boundary); head(sl)
library(sf)
(sl_sf = st_as_sf(sl, coords = c("Var1", "Var2"), crs=4326))
sl_sf$elev = sl_elev
sl_sf
library(tmap)
tm_shape(sl_sf) + tm_dots(col="elev") # plot
```

If one needs to convert an sf object to a raster object, the way to do so is illustrated in the following chunk.

```{r}
ext(sl_sf)
(temp = rast(nrow=200, ncols=100, ext=ext(sl_sf))) # empty raster layer
(sl = rasterize(sl_sf, temp, field="elev", fun=mean)) # Assign values to the empty raster. We average the elevation of the points that go into each pixel.
plot(sl, main="Sri Lanka")
tm_shape(sl) + tm_raster() # plot
```

