# Question 1B: Spatial Regression (20 marks)

Shared with you are two files:

1. “italy.shapefile”, a shapefile consisting geometries of the Italian regions along with the following attributes,

a. REGION: name of the Italian region

b. VAR_UNEMP: variation in unemployment

c. VAR_RGDP: variation in real GDP

2. “it.nb”, a GAL file to generate queen contiguity neighbours of the Italian regions.

```{r}
library(sf); library(spdep); library(spatialreg)
```

Answer the following questions:

a) (5 marks) Load the italy.shapefile and regress VAR_UNEMP on VAR_RGDP. (The inverse relationship between the variation in unemployment and variation in real GDP is known as Okun’s law in the economics literature.)

```{r}
(italy = st_read("italy_okun.shp"))

# OLS
OLS = lm(VAR_UNEMP~VAR_RGDP, data=italy)
```

a. Show the summary of the regression.

```{r}
summary(OLS)
```

b. Give a scatterplot of the variables along with the fitted line.

```{r}
plot(italy$VAR_RGDP,italy$VAR_UNEMP, main="Okun's Law in Italian Regions", 
     ylab="Variation of unemplyment rate", xlab="Variation of real GDP")
abline(OLS,col="Blue")
```

c. Check for homoskedasticity and normality of the regression errors.

```{r}
lmtest::bptest(OLS) # Do not reject H0: homoskedasticity
tseries::jarque.bera.test(OLS$residuals) # Do not reject H0: normal
```

b) (5 marks) Load the GAL file and create a neighbourhood connectivity plot on a map of Italy. Conduct Moran’s tests on the OLS residuals and the model variables to check whether spatial spill-over effects need to be considered in the model.

```{r}
# nb
# (it_nb = poly2nb(italy))
(it_nb = read.gal("italy.gal"))

it_coords = st_point_on_surface(st_geometry(italy))
plot(st_geometry(italy), border="grey60",  main="Queen Contiguity")
plot(it_nb, it_coords, pch=19, cex=0.6, add=T)

(it_lw = nb2listw(it_nb))

# Tests to detect spatial autocorrelation

# in OLS residuals
lm.morantest(OLS, it_lw) # Reject null of CSR
moran.plot(OLS$residuals, it_lw, main="Moran Plot for OLS residuals")

# in other variables
moran.test(italy$VAR_UNEMP, it_lw) # Reject null of CSR
moran.test(italy$VAR_RGDP, it_lw) # Do not reject the null of CSR
```

c) (10 marks) Run the usual spatial regressions and conduct appropriate tests to identify the most suitable spatial regression model.

```{r}
unemp_SAR = spautolm(formula = VAR_UNEMP ~ 1, data=italy, listw=it_lw)
unemp_SLM = lagsarlm(formula = VAR_UNEMP ~ VAR_RGDP, data=italy, listw=it_lw)
unemp_SEM = errorsarlm(formula = VAR_UNEMP ~ VAR_RGDP, data=italy, listw=it_lw)
unemp_SARAR = sacsarlm(formula = VAR_UNEMP ~ VAR_RGDP, data=italy, listw=it_lw)

# tests
lm.LMtests(OLS, listw=it_lw, test="all") # possible SLM

LR1.Sarlm(unemp_SLM) # Reject H0:OLS
LR1.Sarlm(unemp_SARAR) # Reject H0:OLS
LR1.Sarlm(unemp_SEM) # Reject H0:OLS

LR.Sarlm(unemp_SEM,unemp_SARAR) # Reject H0
LR.Sarlm(unemp_SLM,unemp_SARAR) # Reject H0
LR.Sarlm(unemp_SAR,unemp_SLM) # Reject H0

# Overall SARAR model is better suited.
```