# ECON6027: Assignment 1 (50 marks)

Submission format:

1. Original R script or R markdown file

2. HTML/PDF output (knit your script to an html/pdf file for submission)

3. .txt file of the standard deviational ellipses

```{r}
library(sf); library(tmap); library(aspace)
```

### Question 1
a) Load the “cycle_hire” dataset from the package “spData” which is a point dataset of cycle hire points in central London. (5 marks)

```{r}
# Load the dataset
cycle_sf = st_read(system.file("shapes/cycle_hire.geojson", package="spData"))
st_is_longlat(cycle_sf)
```

b) Compute a “standard deviational ellipse” of the cycle hire points. (5 marks)

```{r}
# Apply a projection. I used the British National Grid
# Typically this step is compulsory before you run any calculations.
cycle_sf = cycle_sf %>% st_transform(27700) 
tmap_leaflet(tm_shape(cycle_sf) + tm_dots())

# Standard deviational ellipse (using aspace 4.1.0)
cycle_sde = calc_sde(id=1, points=st_coordinates(cycle_sf))

# change the sde to sf linestring
(cycle_sde_sf = cycle_sde$LOCATIONS %>% st_as_sf(coords = c("x", "y"), crs=27700) %>% 
    st_combine() %>% st_cast("LINESTRING"))
(cycle_sde_sf = st_sf(cycle_sde$ATTRIBUTES, geom=st_geometry(cycle_sde_sf)))
```

c) Plot both hire points and the ellipse on an interactive map using the leaflet driver “OpenStreetMap”. (5 marks)

```{r}
# interactive map
tmap_mode("view")
tm_shape(cycle_sf) + tm_dots() + 
  tm_shape(cycle_sde_sf) + tm_lines(col="red", lwd=2) + 
  tm_layout(main.title = "Cycle Hire Locations in Central London") +
  tm_basemap(server = "OpenStreetMap")
tmap_mode("plot")
```

Don't forget to clear the environment before you proceed.

```{r}
rm(list=ls())
```

### Question 2
The “quakes” data.frame (an inbuilt dataset in R) that gives the locations of 1000 seismic events of MB > 4.0. The events occurred in a cube near Fiji since 1964.

a. Give spatial awareness to this dataset. The coordinate reference system should use the world geodetic system. Generate the head of your dataset. (5 marks)

```{r}
# inspect data
data(quakes); head(quakes)

# main
quakes_sf = st_as_sf(quakes, coords=c("long","lat"))
st_crs(quakes_sf) = 4326 # it is known that the dataset follows the WGS84.
quakes_sf
```

b. Prepare two interactive maps of the recorded “magnitude” and depth of the “location” using “tmap” package. (5 marks)

```{r}
tmap_mode("view")
tm_shape(quakes_sf) + tm_dots(col="mag", title = "magnitude")
tm_shape(quakes_sf) + tm_dots(col="depth")
tmap_mode("plot")
```

c. Compute the (i) mean centre, and (ii) the standard deviational ellipse using “magnitude” attribute as weights. (5 marks)

```{r}
# Descriptive measures cannot be computed using geographic CRS
# Change the CRS from geographic to projected using UTM 
# Fiji 1956 / UTM zone 60S - EPSG:3141.
quakes_sf = st_transform(quakes_sf, crs=3141)
coords = st_coordinates(quakes_sf)
mag = quakes_sf$mag

# (i) mean centre
fiji_mean = calc_mnc(id=2, weighted = T, weights=mag, points=coords)
(fiji_mean_sf = st_as_sf(fiji_mean$ATTRIBUTES, coords = c("CENTRE.x", "CENTRE.y"), crs=3141))

# (ii) SDE
fiji_sde = calc_sde(id=1, weighted = T, weights=mag, points=coords)
(fiji_sde_sf = fiji_sde$LOCATIONS %>% 
    st_as_sf(coords = c("x", "y"), crs=3141) %>% st_combine() %>% 
    st_cast("LINESTRING"))
(fiji_sde_sf = st_sf(fiji_sde$ATTRIBUTES, geom=st_geometry(fiji_sde_sf)))
```

d. Plot the output of (c) on a static map. Give an appropriate title (5 marks)

```{r}
# plot
tm_shape(quakes_sf) + tm_dots(col="mag", title = "magnitude", palette="Blues") + 
  tm_shape(fiji_mean_sf) + tm_squares(col="black") + 
  tm_shape(fiji_sde_sf) + tm_lines(col="black") + 
  tm_layout(main.title="Center and Orientation of Quakes") + 
  tm_add_legend(type="symbol", col="black", labels="Mean Center", shape=15) + 
  tm_add_legend(type="line", col="black", labels="SD Ellipse", lty=1)
# notice how the orientation of the quakes is along the Australian and Pacific plate intersection.
```

Don't forget to clear the environment before you proceed.

```{r}
rm(list=ls())
```

### Question 3
Variables (in order of appearance):

- Planning area name

- Total population (thousands)

- Percentage of population who are free thinkers

- Percentage of population with no literacy

- Percentage of population with tertiary education

- Percentage of population with only primary education

- Percentage of population in white collar jobs

- Total number of households (thousands)

- Average household size by residents

- Percentage of households with a combined income above the median national income

- Percentage of households with a combined income above $20,000

- Percentage of households with a combined income below $1,000

- Percentage of households that are owner occupied

- Percentage of HDBs

```{r}
library(dplyr)
```

a. Load the “sg_house.csv” dataset to R using the read.csv() function. Read the Singapore planning area shapefile from Chapter 1 (shared for your convenience). Combine the two datasets (using an appropriate method). The final object should be an sf object. Your uploaded answer should include a compressed folder of this combined sf object saved as a shapefile. (5 marks)

```{r}
sg_house = read.csv("sg_house.csv")
class(sg_house)

# setwd("C:/Users/sfliu/Dropbox/MSc/Lessons/1 Introduction")
sg_PA = st_read("MP14_PLNG_AREA_WEB_PL.shp") 
table(st_is_valid(sg_PA)) # 8 false entries
# It is always a good idea to check the validity of the object before proceeding. 
# If there are FALSE entries, use st_make_valid to validate the object.
sg_PA = st_make_valid(sg_PA); table(st_is_valid(sg_PA)) 
# next time directly the run the following command.
# sg_PA = st_read("MP14_PLNG_AREA_NO_SEA_PL.shp") %>% st_make_valid()

sg_house_sf = inner_join(sg_house, sg_PA)
class(sg_house_sf) # is it an sf object?
(sg_house_sf = st_as_sf(sg_house_sf))
class(sg_house_sf) # check again...

# Create a shapefile of the object 
st_write(sg_house_sf, "sg_house.shp", delete_layer=T)
```

b. Generate a plot (using tmap) of the attribute, percentage of households above median income given by “PCT_ABOVEMEDIN”. Your map should, (5 marks)

1. Include an appropriate title

2. Use the quartiles as the “breaks”

3. Change the line type to “dotted”.

```{r}
summary(sg_house_sf$PCT_ABOVEMEDIN) # check quartiles
bins = c(34.52, 43.04, 50, 58.95, 73.22)
tm_shape(sg_house_sf) + tm_polygons(col="PCT_ABOVEMEDIN", breaks=bins, lty=3) +
  tm_layout(main.title = "Percentage of Households Above Median Income, SG, 2015")
# plot(sg_house_sf["PCT_ABOVEMEDIN"], breaks=bins, lty=3, 
#     main="Percentage of Households Above Median Income, Singapore, 2015")
```

c. Extract the central region and create a new spatial object. Create a subset of planning areas in the central region that has more than half the households listed as HDBs. (5 marks)

```{r}
(central_sg = sg_house_sf[sg_house_sf$REGION_C == "CR", ]) # Extract central region
(central_hdb = subset(central_sg, PCT_HDB > 50))

# Codes may generally differ and that is OK!
```