---
title: "ECON6027 project (Group 4)"
output: html_document
date: "2024-10-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(aspace, quietly = T)
library(sf, quietly = T)
library(tmap, quietly = T)
library(spatstat, quietly = T)
library(ggplot2, quietly = T)
library(dplyr, quietly = T)
library(geojsonsf, quietly = T)
library(geojsonR, quietly = T)
```

We start by loading the shapefile of Singapore planning areas from [data.gov.sg](https://data.gov.sg/datasets?query=master+plan+2014&page=1&resultId=d_6d6fd6505f23d7fb90dec567afd555bb) and the geojson file of MRT and LRT exits from [data.gov.sg](https://data.gov.sg/datasets?query=mrt&page=1&resultId=d_b39d3a0871985372d7e1637193335da5).

```{r}
sg_regions = st_read("MP14_PLNG_AREA_WEB_PL.shp")
sg_regions = st_transform(sg_regions, crs=3414)
table(st_is_valid(sg_regions)) # there are 5 false entries
sg_regions = st_make_valid(sg_regions)
table(st_is_valid(sg_regions)) # now, all 55 entries are valid

mrt_lrt_exits = geojson_sf("LTAMRTStationExitGEOJSON.geojson")
mrt_lrt_exits_sf = st_transform(mrt_lrt_exits, crs=3414)
table(st_is_valid(mrt_lrt_exits_sf)) # all entries are valid
st_crs(mrt_lrt_exits_sf)$proj4string
isTRUE(all.equal(st_crs(sg_regions), st_crs(mrt_lrt_exits_sf))) #TRUE, we can proceed
table(is.na(mrt_lrt_exits_sf$geometry)) # no station exits with empty coordinates 


sg_map_bg = tm_shape(sg_regions) +
  tm_polygons(alpha=0.5, border.col="black", legend.show=F) +
  tm_fill(col = "white") + 
  tm_scale_bar(position = c("right", "top"), width = 0.25, text.size = 0.4) +
  tm_compass(position = c("left", "top"), size =0.8, text.size = 0.7)
```

## 1. Spatial descriptive summary measures
Computations using the aspace package functions can only be carried out on cartesian coordinates given in matrix form. Thus, we will extract the coordinates of MRT and LRT exits.

```{r}
mrt_coords = st_coordinates(mrt_lrt_exits_sf)
mrt_coords = st_zm(mrt_coords, drop = T, what = "ZM")
class(mrt_coords)
```

Measures of central tendency: We compute the mean and the median and assign them to a list object for later use.

```{r}
mean_mrt = calc_mnc(id=1, points=mrt_coords)
mean_mrt$LOCATIONS
median_mrt = calc_mdc(id=2, points=mrt_coords)
median_mrt$LOCATIONS
```

In the following chunk, we compute the standard deviation distance and ellipse of mrt and lrt exits in Singapore. 

```{r}
sdd_mrt = calc_sdd(id=3, calccentre = T, weighted=F, points = mrt_coords)
sde_mrt = calc_sde(id=4, calccentre=T, weighted=F, points = mrt_coords)

mrt_sdd_line = sdd_mrt$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs=st_crs(mrt_lrt_exits_sf)) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")
mrt_std_dist = st_sf(sdd_mrt$ATTRIBUTES, geom=st_geometry(mrt_sdd_line))

mrt_sde_line = sde_mrt$LOCATIONS %>% 
  st_as_sf(coords = c("x", "y"), crs=st_crs(mrt_lrt_exits_sf)) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")
mrt_std_ellps = st_sf(sde_mrt$ATTRIBUTES, geom=st_geometry(mrt_sde_line))

mean_mrt$LOCATIONS; median_mrt$LOCATIONS #identify mean and median location, ID 1 is mean and ID 2 is median
mean_median = rbind(mean_mrt$LOCATIONS, median_mrt$LOCATIONS) %>% data.frame() %>% 
  st_as_sf(coords=c("x","y"), crs=st_crs(mrt_lrt_exits_sf))
mean_median$id = c("Spatial mean","Spatial median")

sg_map_bg + 
  tm_shape(mrt_lrt_exits_sf) + tm_dots(size=0.1, col= "blue") +
  tm_shape(mrt_std_ellps) + tm_lines(lwd=2) + 
  tm_shape(mrt_std_dist) + tm_lines(lwd=2, lty=2) + 
  tm_shape(mean_median) + tm_dots(size=0.1, shape="id", col="red") +  
  tm_add_legend(type="line", labels="SDD", lty=2, ) + 
  tm_add_legend(type="line", labels="SDE", lty=1)
```

## 2. Quadrat count analysis

```{r}
mrt.ppp = as.ppp(mrt_lrt_exits_sf$geometry) #transform sf object to ppp
plot(mrt.ppp, pch=20, size=0.5)
st_bbox(mrt_lrt_exits_sf)
```

```{r}
sg = st_union(sg_regions) 
sg.owin = as.owin(sg)
class(sg.owin)
Window(mrt.ppp) = sg.owin 
mrt.ppp # 563 points, no points out of the boundary

table(duplicated(mrt.ppp)) # no duplicated entries, we can proceed

qc_mrt = quadratcount(mrt.ppp, nx=10, ny=10)
class(qc_mrt)
plot(qc_mrt)
```

Chi2 (goodness of fit) test

```{r}
quad.test_mrt = quadrat.test(mrt.ppp, nx=10, ny=10)
```

Test outcome: p-value < 2.2e-16, reject the null at 5% level.

## 3. Kernel density estimation

```{r}
plot(density(mrt.ppp, sigma=500))
plot(density(mrt.ppp, sigma=1000))

mrt.bw1 = bw.diggle(mrt.ppp)
mrt.kde1 = density(mrt.ppp, sigma=mrt.bw1)
plot(mrt.kde1)

mrt.bw2 = bw.ppl(mrt.ppp)
mrt.kde2 = density(mrt.ppp, sigma=mrt.bw2)
plot(mrt.kde2)

mrt.bw3 = bw.scott(mrt.ppp)
mrt.kde3 = density(mrt.ppp, sigma=mrt.bw3)
plot(mrt.kde3)

mrt.bw4 = bw.CvL(mrt.ppp)
mrt.kde4 = density(mrt.ppp, sigma=mrt.bw4)
plot(mrt.kde4)

mrt.bw5 = bw.frac(mrt.ppp)
mrt.kde5 = density(mrt.ppp, sigma=mrt.bw5)
plot(mrt.kde5)
```

## 4. Nearest neighbour distance analysis
```{r}
clarkevans(mrt.ppp, correction="none") #calculated index values less than 1 (r=0.1399807)
clarkevans.test(mrt.ppp, correction="none", alternative="less")

mrt.nnd = nndist(mrt.ppp, k=1)
summary(mrt.nnd)
hist(mrt.nnd)
class(mrt.nnd)
mrt.nnd = as.matrix(mrt.nnd)
```

## 5. K-function


## 6. Real-world analysis
Patching in population density by planning areas from [singstat.gov.sg](https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data). An area of consideration could be whether there are enough MRT/LRT in each planning area for residents of Singapore. This can be seen by patching a layer of population density onto the Singapore planning area map with the MRT/LRT exit points.

```{r}
population_pa = read.csv("residentpopulation_2024.csv")
population_sf = inner_join(population_pa, sg_regions)
population_sf = st_as_sf(population_sf) 
population_sf= st_transform(population_sf, 3414)
isTRUE(all.equal(st_crs(sg_regions), st_crs(population_sf))) # TRUE
population_sf$resident_2024_int = as.numeric(gsub(",", "", population_sf$resident_2024))

population_density = 
  tm_shape(population_sf) + 
  tm_borders(col="black", lwd=1, lty=1) +
  tm_fill("resident_2024_int", lty=3, palette = "Blues") +
  tm_scale_bar(position = c("right", "bottom")) +
  tm_compass(position = c("left", "top"), size =0.8, text.size = 0.7) + 
  tm_layout(main.title = "Population density map", legend.outside = T) + tm_text("PLN_AREA_N", size = 0.25, clustering = T)

mrt = population_density + tm_shape(mrt_lrt_exits_sf) + 
  tm_dots(col = "red", size=0.01) +
  tm_layout(main.title = "MRT and LRT exits against population density map", legend.outside = T, main.title.size = 0.7)
mrt
```

From the map, we can see that there is a clustering of MRT/LRT stations at the central region with areas such as Newton, Orchard, River Valley, Downtown Core, and Outram despite the low population density.