# Question 2: Geospatial Data (30 marks)

Topo.Rdata provides a data frame with 52 rows and 3 columns, of topographic height in feet within a 310 feet square. The objective of this question is to recreate the topology using geospatial interpolation techniques.

```{r}
library(sf); library(tmap); library(gstat)
```
 
a. Give spatial awareness to the dataset. Give a basic plot of your dataset. (3 marks)

```{r}
tmap_options(show.warnings=F, show.messages=F)

library(MASS)
data(topo) # The topo data frame has 52 rows and 3 columns, of topographic heights within a 310 feet square.
head(topo)

# save(topo, file="topo.Rdata")

(topo = st_as_sf(topo, coords=c("x", "y")))

plot(topo, main="Height, ft")
```

b. Re-create the Voronoi Tesselation given in Figure 3.1. (5 marks)

```{r}
topo_sp = as(topo,"Spatial")
voro_sp = dismo::voronoi(topo_sp)

(voro_sf = voro_sp %>% st_as_sf())

(voro_plot = tm_shape(voro_sf) + tm_polygons(col="z", palette="YlGn", title="Height, ft") + 
    tm_shape(topo) + tm_dots() + tm_layout(main.title='Voronoi Tesselation of Heights'))
```

c. Create a square grid of 10,000 unsampled points for the purpose of interpolation. Give a plot of this grid points. (3 marks)

```{r}
bbox = pgirmess::bbox2sf(bbox=st_bbox(topo))
(grid = st_sample(bbox, size=10000, type="regular"))

plot(grid)
```

d. Reconstruct the topology using the inverse distance weighting method where the “idp=1” and “idp=3”. Which idp is more appropriate? (5 marks) 

```{r}
# generally not suitable for elevation data

# idp = 1
(idw1 = idw(formula=z~1, locations=topo, newdata=grid, idp = 1))

(idw_plot1 = tm_shape(idw1) + 
    tm_dots(col="var1.pred", palette="YlGn", title="Height, ft", shape=15) + 
    tm_layout(main.title="IDW interpolation for height, idp=1"))

# idp = 3
(idw3 = idw(formula=z~1, locations=topo, newdata=grid, idp = 3))

(idw_plot3 = tm_shape(idw3) + 
    tm_dots(col="var1.pred", palette="YlGn", title="Height, ft", shape=15) + 
    tm_layout(main.title="IDW interpolation for height, idp=3"))

tmap_arrange(idw_plot1, idw_plot3, nrow=1)

# idp=3 is better than idp=1 when comparing against the Voronoi tessellation
```

e. Give an ordinary kriging (OK) estimate of the topology by selecting the most suitable parametric model of the semi-variogram. When selecting the most suitable parametric model, consider Stein, Gaussian, Bessel and linear fits and choose the best fit. Your output must include, (10 marks)

```{r}
(autovgm = automap::autofitVariogram(z~1, topo_sp)) # Stein model selected

svgm = variogram(z~1, topo)
plot(svgm, main="Empirical semi-variogram") # make initial guesses here or use autofit estimates

(fvgm.s = fit.variogram(svgm, vgm(psill=4000, model="Ste", range=3), warn.if.neg=T)) # Stein does not work
(fvgm.g = fit.variogram(svgm, vgm(psill=4000, model="Gau", range=3), warn.if.neg=T)) # Gau
(fvgm.b = fit.variogram(svgm, vgm(psill=4000, model="Bes", range=3), warn.if.neg=T)) # Bes
(fvgm.l = fit.variogram(svgm, vgm(psill=4000, model="Lin", range=3), warn.if.neg=T)) # Lin

plot1 = plot(svgm, model=fvgm.s, main="Fitted semi-variogram, Stein's model")
plot2 = plot(svgm, model=fvgm.g, main="Fitted semi-variogram, Gaussian model")
plot3 = plot(svgm, model=fvgm.b, main="Fitted semi-variogram, bessel model")
plot4 = plot(svgm, model=fvgm.l, main="Fitted semi-variogram, linear model")
```

1. Model selected with the fitted semi-variogram plot.

```{r}
gridExtra::grid.arrange(plot1, plot2, plot3, plot4, nrow = 2)
# select Gaussian/Bessel model
```

2. Summary of the OK estimates.

```{r}
ok = krige(z~1, locations=topo, newdata=grid, model=fvgm.b)
summary(ok)
# negative variance estimates!

ok = krige(z~1, locations=topo, newdata=grid, model=fvgm.g)
summary(ok)
# negative predictions are valid for topology
# Also, prediction variance is smaller for Gaussian
```

3. Plot of the reconstructed topology (breaks you use for this plot must be comparable to the Voronoi Tessellation).

```{r}
(ok_plot = tm_shape(ok) + 
  tm_dots(col="var1.pred", palette="YlGn", breaks=c(-Inf,700,750,800,850,900,950,Inf), title="Height, ft", shape=15) + 
  tm_layout(main.title="OK estimate for height"))

# varok_plot = tm_shape(ok) + 
#   tm_dots(col="var2.var", palette="OrRd", title="Estaimted variance", shape=15) + 
#   tm_shape(topo) + tm_dots() + 
#   tm_layout(main.title="Variance of the OK height estiamte")
# tmap_arrange(ok_plot, varok_plot)
```

f. Give a plot of the three different reconstructed maps: Voronoi tessellation, best IDW estimate, and OK estimate. (2 marks)

```{r}
tmap_arrange(ok_plot, idw_plot3, voro_plot)
```

g. For the kind of data given in this question, which method of interpolation is most suitable and why? (2 marks)

- The most suitable technique is kriging. The prediction variance is very small throughout the entire map.
- IWD is not suitable since we are reconstructing elevation and IDW is not sensitive to areas of peaks or pits and thus lead to undesirable results.

```{r}
tmap_options_reset()
```